\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-19 Documents/M1/Astronum_2018/sections/Numerical.tex
Buckup the latest changes on M1/Astronum_2018

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution, adopted from \cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain with $D=\{x:x\in[-3,3]\}$ and an initial condition with $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ is used.
The analytic solution is given by $\cJ = \cJ_{0} \times e^{-\f{\pi^2 t}{27\sigma_{\Scatt}}}$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the error versus $N$.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSPRK2 behaves as good as SSP2332 in the diffusion region but requires 1/3 less implicit solvers.
\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
%  \begin{tabular}{cc}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
%  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines on the left plot reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line on the right plot reference lines proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more ``realistic'' test: two-dimensional neutrino transport with emission, absorption, and elastic scattering through a stationary background.
This is designed to test the realizability-preserving properties of the PD-ARS schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities we used.
This test is constructed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.
We take 10 energy groups from 0~MeV to 300~MeV and initialize the neutrino number density $\cJ$ and flux density $\bcH$ both at zero(arbitrarily small).
For the realizability-preserving test, we employ the CB closure with PD-ARS schemes, SSP2332, a scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and a scheme proposed by McClarren et al \cite{mcclarren_etal_2008} are all applied.
Only PD-ARS schemes produce realizability-preserving results and evolve the model to a steady state.
The other schemes crashed immediately after several steps.
The realizability-preserving plot and the evolve process of PD-ARS schemes are given in Figure~\ref{fig:NeutrinoStationaryTestEvolve}.

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Realizability (left column), the number density $\cJ$ versus radius (center column), and the flux factor $|\bcH|/\cJ$ versus radius (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, the light blue area highlights the realizability domain, and the black lines define its boundary. Each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

>>>>

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution, adopted from \cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain with $D=\{x:x\in[-3,3]\}$ and an initial condition with $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ is used.
The analytic solution is given by $\cJ = \cJ_{0} \times e^{-\f{\pi^2 t}{27\sigma_{\Scatt}}}$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the error versus $N$.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSPRK2 behaves as good as SSP2332 in the diffusion region but requires 1/3 less implicit solvers.

\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
%  \begin{tabular}{cc}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
%  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines on the left plot reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line on the right plot reference lines proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more ``realistic'' test: two-dimensional neutrino transport with emission, absorption, and elastic scattering through a stationary background.
This is designed to test the realizability-preserving properties of the PD-ARS schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities we used.
This test is constructed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.
We take 10 energy groups from 0~MeV to 300~MeV and initialize the neutrino number density $\cJ$ and flux density $\bcH$ \rc{both at zero(arbitrarily small) give the number}.
For the realizability-preserving test, we employ the CB closure with PD-ARS schemes, SSP2332, a scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and a scheme proposed by McClarren et al \cite{mcclarren_etal_2008} are all applied.
Only PD-ARS schemes produce realizability-preserving results and evolve the model to a steady state.
The other schemes crashed immediately after several steps.
The realizability-preserving plot and the evolve process of PD-ARS schemes are given in Figure~\ref{fig:NeutrinoStationaryTestEvolve}. 
\rc{More details of what's plotted.}

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Realizability (left column) \rc{Explain the realizability difference(light blue and light red)}, the number density $\cJ$ versus radius (center column), and the flux factor $|\bcH|/\cJ$ versus radius (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, the light blue area highlights the realizable domain, and the black lines define its boundary. Each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

### 2018-10-19 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Buckup the latest changes on M1/Astronum_2018

\subsection{Standard Implicit-Explicit Scheme}
Treating $\vect{\cT}$ explicitly and $\f{1}{\tau}\,\vect{\cQ}$ implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where  $(\tilde{a}_{ij})$ and $(a_{ij})$ are coefficients of the $i$-th stage and they are the elements of matrices $\tilde{A}$ and $A$, respectively.
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weight vectors.
These coefficients and weights have to satisfy certain condition for an accurate and consistent scheme.
For instance, to have second-order temporal accuracy, the following conditions are required:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j}^{s}a_{ij}$.

\subsection{A Constraint-Preserving Implicit-Explicit Scheme}
Beyond the accuracy requirement, a convex-invariant IMEX scheme has extra constraints to satisfy.
The problem left is to find the additional constraints on $(\tilde{a}_{ij})$, $(a_{ij})$, $\tilde{w}_{i}$, and $w_{i}$.
Following Hu et al\cite{hu_etal_2018}, the stage values in Eq.~\eqref{imexStages} can be rewritten as
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
with $c_{ij}$ and $\hat{c}_{ij}$ be given by $a_{ij}$ and $\tilde{a}_{ij}$.

It can be proven that the conditions we need for a simple convex-invariant IMEX scheme are: \cite{chu_etal_2018}

>>>>

\subsection{Standard Implicit-Explicit Scheme}
Treating $\vect{\cT}$ explicitly and $\f{1}{\tau}\,\vect{\cQ}$ implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}: 
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where  $(\tilde{a}_{ij})$ and $(a_{ij})$ are coefficients of the $i$-th stage and they are the elements of matrices $\tilde{A}$ and $A$, respectively.
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weights in the assembly step in Eq.~\eqref{imexIntermediate}.
These coefficients and weights have to satisfy certain condition for an accurate and consistent scheme.
For instance, to have second-order temporal accuracy, the following conditions are required:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j}^{s}a_{ij}$.

\subsection{A Constraint-Preserving Implicit-Explicit Scheme}
Beyond the accuracy requirement, a convex-invariant IMEX scheme must satisfy additional constraints.
Our idea is finding the constraints on $a_{ij}$, $\tilde{a}_{ij}$, $\tilde{w}_{i}$, and $w_{i}$ that enable each $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ be convex combination of realizable moments:

Following Hu et al\cite{hu_etal_2018}, the stage values in Eq.~\eqref{imexStages} can be rewritten as
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
with $c_{ij}$ and $\hat{c}_{ij}$ be some expression of $a_{ij}$ and $\tilde{a}_{ij}$.
By requiring the IMEX scheme be globally stiffly accurate (GSA), i.e. $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i}$ for $i=1,\ldots,s$, $\vect{u}^{n+1} = \vect{u}^{(s)}$.
Using those Lemma proved in \cite{chu_etal_2018}, we can find those constraints that $c_{ij}$ and $\hat{c}_{ij}$ (or $a_{ij}$ and $\tilde{a}_{ij}$) need to satisfy to ensure $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ be realizable.

\rc{Diffusion accurate detail...}

\rc{Only ARS and define ARS...}

Here is list of our constraints \cite{chu_etal_2018}:

### 2018-10-19 Documents/M1/Astronum_2018/sections/TwoMoments.tex
Buckup the latest changes on M1/Astronum_2018

\section{Two-Moment Model}\label{se:Two-MomentModel}

@@

The inequalities in Eq.~\eqref{eq:eddingtonFactorBounds} deserve more attention than has been given.
They have equal importance relative to the inequalities in Eq.~\eqref{eq:MomentsBounds} in maintain consistency with respect to Fermi-Dirac statistics.

The Eddington factor given by the algebraic closures discussed in \cite{murchikova_etal_2017} are examined in this paper as examples.
We consider various occupancies and plot at $\cJ = 0.1$ and $\cJ = 0.9$ as examples of low occupancy and high occupancy, respectively, in Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure}.  
As the figure shows, not all algebraic closures satisfy the Eddington factor bounds, Eq.~\eqref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2\cite{janka_1992} may work fine in the neutrino low occupancy limit, but not for high occupancies; Janka 1\cite{janka_1991} can be in error in both cases.
The only algebraic closure that remains within the Eddington factor bounds among those plotted is Cernohorsky and Bludman's \cite{cernohorskyBludman_1994}.

Though Levermore closure and Minerbo closure don't conserve the bounds in the high occupancy limit, they are widely used: {O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} use Minerbo closure, and Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016} use Levermore clsoure.

>>>>

\section{Two-Moment Model}
\label{se:Two-MomentModel}

@@

The inequalities in Eq.~\eqref{eq:eddingtonFactorBounds} deserve more attention than has been given.
They have equal importance relative to the inequalities in Eq.~\eqref{eq:MomentsBounds} in maintain consistency with respect to Fermi-Dirac statistics.
The Eddington factor given by the algebraic closures discussed in \cite{murchikova_etal_2017} are examined in this paper as examples.
We consider various occupancies and plot $\cJ = 0.1$ and $\cJ = 0.9$ as examples for low and high occupancy, respectively, in Fig.~\eqref{fig:EddingtonFactorsWithDifferentClosure}.
As the figure shows, not all algebraic closures satisfy the Eddington factor bounds, Eq.~\eqref{eq:eddingtonFactorBounds}: Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 2\cite{janka_1992} may work fine in the neutrino low occupancy limit, but not for high occupancies; Janka 1\cite{janka_1991} can be in error in both cases.
The only algebraic closure that remains within the Eddington factor bounds among those plotted is Cernohorsky and Bludman's \cite{cernohorskyBludman_1994}.
Though Levermore closure and Minerbo closure don't conserve the bounds in the high occupancy limit, they are widely used: {O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} use Minerbo closure, and Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016} use Levermore clsoure.
\rc{Add more discussion.}

### 2018-10-19 Documents/M1/Astronum_2018/sections/intro.tex
Buckup the latest changes on M1/Astronum_2018

Core-collapse supernovae (CCSNe) are the explosions of massive stars that end their lives.
CCSNe are a dominant source of heavy elements and play an important role in many astrophysical phenomena, such as neutron star and black hole formation.  
Furthermore, these explosions occur at energies and densities relevant to address fundamental questions in nuclear, particle, and gravitational physics. 
A solid theoretical framework for the CCSN explosion mechanism may help answer important questions in fundamental physics.

@@

However, solving the multi-dimensional Boltzmann transport equation is expensive.
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted.
In the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$: the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the transport term in the Boltzmann transport equation is the dot product of the velocity vector and the gradient of the phase-space distribution function.
Integrating this term with unit direction vector as the weight introduces the second angular moment $\bcK$ to the moment equations.
Knowledge of $\bcK$ is needed to close the equation system.
Therefore, an algebraic closure that gives an approximate value of $\bcK$ given $\cJ$ and $\bcH$ is needed.
The two-moment method has been widely applied with different algebraic closures, such as the Minerbo\cite{minerbo_1978} closure(e.g.~{O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan and et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} ) and the Levermore\cite{levermore_1984} closure(e.g.~Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016}).

Applying the two-moment method alone does not guarantee an affordable solution.
To be precise, neutrinos interact with the background, and the characteristic time scale of these interactions, which can be $\sim\mathcal{O}(10^{-13})$~second, is short compared to the duration of the CCSN explosion, which $\sim\mathcal{O}(1)$~second.  
This makes the time step needed by an explicit time integrator prohibitively small. 
On the other hand, solving the moment equations fully implicitly requires inverting globally many band-structured matrices whose sizes depend on the spatial discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require fewer time steps comparing with fully explicit method, and the calculation for each step is easily parallelizable.  
For a non-relativistic system where the propagating speed of gravity wave and fluid is much slower than the light speed, fully implicit method is a better choice for neutrino transport in CCSN simulation.
But for a relativistic circumstance that we have, where the gravity, fluid, and neutrino have a relativistic propagating speed, IMEX method
balances the efficiency and cost.

>>>>

Core-collapse supernovae (CCSNe) are the explosions of massive stars that end their lives.
CCSNe are a dominant source of heavy elements and play an important role in many astrophysical phenomena, such as neutron star and black hole formation.  
Furthermore, these explosions occur at energies and densities relevant to address fundamental questions in nuclear, particle, and gravitational physics. 
A solid theoretical framework for the CCSN explosion mechanism may help answer important questions in fundamental physics.\rc{[references]}

@@

However, solving the multi-dimensional Boltzmann transport equation is expensive.\rc{[references]}
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted.\rc{[references]}
In the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$: the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the transport term in the Boltzmann transport equation is the dot product of the velocity vector and the gradient of the phase-space distribution function.
Integrating this term with unit direction vector as the weight introduces the second angular moment $\bcK$ to the moment equations.
Knowledge of $\bcK$ is needed to close the equation system.
Therefore, an algebraic closure that gives an approximate value of $\bcK$ given $\cJ$ and $\bcH$ is needed.
The two-moment method has been widely applied with different algebraic closures, such as the Minerbo\cite{minerbo_1978} closure(e.g.~{O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan and et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} ) and the Levermore\cite{levermore_1984} closure(e.g.~Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016}).

Applying the two-moment method alone does not guarantee an affordable solution.
To be precise, neutrinos interact with the background, and the characteristic time scale of these interactions, which can be $\sim\mathcal{O}(10^{-13})$~second, is short compared to the duration of the CCSN explosion, which $\sim\mathcal{O}(1)$~second.  
This makes the time step needed by an explicit time integrator prohibitively small. 
On the other hand, solving the moment equations fully implicitly requires inverting globally many band-structured matrices whose sizes depend on the spatial discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require fewer time steps comparing with fully explicit method, and the calculation for each step is easily parallelizable.  
For a non-relativistic system where the propagating speed of gravity wave and fluid is much slower than the light speed, fully implicit method is a better choice for neutrino transport in CCSN simulation.\rc{[references]}
But for a relativistic circumstance that we have, where the gravity, fluid, and neutrino have a relativistic propagating speed, IMEX method
balances the efficiency and cost.
